# Intro to Data Science HW 7
##### Copyright Jeffrey Stanton, Jeffrey Saltz, and Jasmina Tacheva


```{r}
# Enter your name here: Nick Videtti
```

### Attribution statement: (choose only one and delete the rest)


```{r}
# 1. I did this homework by myself, with help from the book and the professor.
```

The chapter on **linear models** (“Lining Up Our Models”) introduces **linear predictive modeling** using the tool known as **multiple regression**. The term “multiple regression” has an odd history, dating back to an early scientific observation of a phenomenon called **“regression to the mean.”** These days, multiple regression is just an interesting name for using **linear modeling** to assess the **connection between one or more predictor variables and an outcome variable**. 


<br>In this exercise, you will **predict Ozone air levels from three predictors**.

A.	We will be using the **airquality** data set available in R. Copy it into a dataframe called **air** and use the appropriate functions to **summarize the data**. 


```{r}
air <- airquality
str(air)
#There are 153 rows and 6 columns (Ozone, Solar.R, Wind, Temp, Month, Day) in these data. Five variables are integers (Ozone, Solar.R, Temp, Month, Day) and one is a numeric (Wind).

summary(air)
#Relatively wide range for everything, but hard to confirm that without knowing more about these variables. This is maybe not as important to consider when looking at Day or Month, but what is important for those is that these data appear to span over five months, at least one of which has 31 days in it. It is probably safe to assume that these months were consecutive and that they were May through September.
```

B.	In the analysis that follows, **Ozone** will be considered as the **outcome variable**, and **Solar.R**, **Wind**, and **Temp** as the **predictors**. Add a comment to briefly explain the outcome and predictor variables in the dataframe using **?airquality**.


```{r}
?airquality

#Outcome Variable:
#Ozone - Mean ozone in parts per billion from 1300 to 1500 hours at Roosevelt Island

#Predictor Variables:
#Solar.R - Solar radiation in Langleys in the frequency band 4000–7700 Angstroms from 0800 to 1200 hours at Central Park
#Wind -  Average wind speed in miles per hour at 0700 and 1000 hours at LaGuardia Airport
#Temp - Maximum daily temperature in degrees Fahrenheit at La Guardia Airport
```

C.	Inspect the outcome and predictor variables – are there any missing values? Show the code you used to check for that.


```{r}
#Outcome Variable
summary(air$Ozone)
#OR
nrow(air[is.na(air$Ozone),])
#Ozone has 37 missing values

summary(air$Solar.R)
#OR
nrow(air[is.na(air$Solar.R),])
#Solar.R has 7 missing values

summary(air$Wind)
#OR
nrow(air[is.na(air$Wind),])
#Solar.R has no missing values

summary(air$Temp)
#OR
nrow(air[is.na(air$Temp),])
#Solar.R has no missing values
```

D.	Use the **na_interpolation()** function from the **imputeTS package** (remember this was used in a previous HW) to fill in the missing values in each of the 4 columns. Make sure there are no more missing values using the commands from Step C.


```{r}
library(imputeTS)

#We will need to fill in missing values in Ozone and Solar.R
air$Ozone <- na_interpolation(air$Ozone)
air$Solar.R <- na_interpolation(air$Solar.R)

#Check to see if this worked:
#Ozone
summary(air$Ozone)
#OR
nrow(air[is.na(air$Ozone),])

#Solar.R
summary(air$Solar.R)
#OR
nrow(air[is.na(air$Solar.R),])

#No more missing values!!
```

E.	Create **3 bivariate scatterplots (X-Y) plots** (using ggplot), for each of the predictors with the outcome. **Hint:** In each case, put **Ozone on the Y-axis**, and a **predictor on the X-axis**. Add a comment to each, describing the plot and explaining whether there appears to be a **linear relationship** between the outcome variable and the respective predictor.


```{r}
library(tidyverse)

#Solar.R on x-axis, Ozone on y-axis
ggplot(air,aes(x = Solar.R, y = Ozone)) + geom_point()
#There does appear to possibly be a linear relationship in our data between Solar.R and Ozone.

#Wind on x-axis, Ozone on y-axis
ggplot(air,aes(x = Wind, y = Ozone)) + geom_point()
#There does appear to possibly be a linear relationship in our data between Wind and Ozone.

#Temp on x-axis, Ozone on y-axis
ggplot(air,aes(x = Temp, y = Ozone)) + geom_point()
#There does appear to possibly be a linear relationship in our data between Temp and Ozone.
```

F.	Next, create a **simple regression model** predicting **Ozone based on Wind**, using the **lm( )** command. In a comment, report the **coefficient** (aka **slope** or **beta weight**) of **Wind** in the regression output and, **if it is statistically significant**, **interpret it** with respect to **Ozone**. Report the **adjusted R-squared** of the model and try to explain what it means. 


```{r}
PartF <- lm(Ozone ~ Wind, data = air)
PartF
summary(PartF)
#Coefficient of Wind is -4.5925. This is statistically significant. This can be interpreted as "for every mile per hour increase in the average wind speed at 7AM and 10AM at LaGuardia Airport, the mean ozone from 1PM to 3PM at Roosevelt Island is expected to decrease by 4.5925 parts per billion.
#The adjusted R-squared is 0.2527, or 25.27%. This means that 25.27% of the change in Ozone in our data can be explained by the change in Wind.
```

G.	Create a **multiple regression model** predicting **Ozone** based on **Solar.R**, **Wind**, and **Temp**.<br> **Make sure to include all three predictors in one model – NOT three different models each with one predictor.**


```{r}
PartG <- lm(Ozone ~ Solar.R + Wind + Temp, data = air)
PartG
summary(PartG)
```

H.	Report the **adjusted R-Squared** in a comment – how does it compare to the adjusted R-squared from Step F? Is this better or worse? Which of the predictors are **statistically significant** in the model? In a comment, report the coefficient of each predictor that is statistically significant. Do not report the coefficients for predictors that are not significant.


```{r}
#The adjusted R-squared is 0.4207, or 42.07%. This is much higher than our adjusted R-squared from Part F of 0.2527 or 25.27%. This means that this model is better than the model from Part F, as this model can describe 42.07% of the change in Ozone, 16.8% more than that of Part F.
#Wind and Temp are statistically significant, but Solar.R is not.
#Coefficient of Wind is -2.69669.
#Coefficient of Temp is 1.53072.
```

I.	Create a one-row data frame like this: 


```{r}
predDF <- data.frame(Solar.R=290, Wind=13, Temp=61)
```

 and use it with the **predict( )** function to predict the **expected value of Ozone**:


```{r}
predict(PartG,predDF)

#Based on our multiple regression model from Part G, when Solar.R is 290, Wind is 13, and Temp is 61, Ozone is expected to be 10.9464.
```

J.	Create an additional **multiple regression model**, with **Temp** as the **outcome variable**, and the other **3 variables** as the **predictors**. 

Review the quality of the model by commenting on its **adjusted R-Squared**.  


```{r}
PartJ <- lm(Temp ~ Ozone + Solar.R + Wind, data = air)
PartJ
summary(PartJ)
#With an adjusted R-squared of 0.403, this model is pretty good relative to the other two we have created in this assignment. This is much better than the simple regression model from Part F (adjusted R-squared of 0.2527), and is pretty close but not quite as good as the multiple regression model from Part G (adjusted R-squared of 0.4207).
```
